The \project{unRAVE} catalog of stellar atmospheric parameters and\\ chemical abundances for chemo-dynamic studies in the \gaia\ era

### 2016-09-09 article/unrave.tex
Update figures and text to gold standard sect

\begin{abstract}
The Milky Way is a powerful laboratory for understanding galaxy formation,
as it can provide orbits, ages, atmospheric parameters, as well as chemical 
abundances for vast sets of individual stars.  These inferences require 
both astrometric and spectroscopic data.  Indeed, in order to be maximally
useful for chemo-dynamic studies, the Tycho-Gaia Astrometric Solution 
(\tgas) sample in the first \gaia\ data release requires a spectroscopic 
complement that includes radial velocities, stellar atmospheric parameters, and 
elemental abundances.  Among existing spectroscopic samples, the RAdial 
Velocity Experiment (\rave) survey has the largest overlap with \tgas: 
$\approx$250,000 stars.  We present a data-driven re-analysis of \rave\ 
spectra using \thecannon, and we provide more precise and accurate stellar
parameters and abundances than available in previous \rave\ data releases. 
We constructed our model using high-fidelity \apogee\ stellar parameters 
and abundances for red giant branch stars that overlap with \rave, and 
stellar parameters from the \project{K2}/\project{Ecliptic Pole Input 
Catalog} for main-sequence and sub-giant branch stars.  We derive, and 
validate, improved effective temperature $\teff$, surface gravity $\logg$, 
and chemical abundances of up to seven elements (O, Mg, Al, Si, Ca, Fe, Ni) 
for \Nstarsqc\ stars.  In total we report \Nabundances\ elemental 
abundances, with a typical precision of 0.07~dex.  The \unrave\ catalog 
presented here, in conjunction with \tgas, represents the most powerful 
data set for chemo-dynamic analysis at the dawn of broadly available 
\gaia\ data.
\end{abstract}

@@

\begin{itemize}
\item We assume that any fibre- and time-dependent variations in spectral
resolution in the \rave\ spectra are negligible.
\item The \rave\ noise variances are approximately correct, independent between
pixels, and normally distributed.
\item We assume that the normalization procedure employed by the \rave\ pipeline
is invariant with respect to the labels we seek to measure (e.g., $\teff$, $\logg$,
or [Fe/H]), and invariant with respect to the S/N ratio.  In other words, we assume
that the normalization procedure does not produce different results for high S/N
spectra compared to low S/N spectra, nor does the normalization procedure vary 
non-linearly with respect to stellar parameters (e.g., [Fe/H]).
\item We assume that stars with similar labels ($\teff$, $\logg$, and abundances)
have similar spectra.
\item A stellar spectrum is a smooth function of the label values for that star,
and we assume that the function is smooth enough within a sub-space of the labels
(e.g., the giant branch or the main-sequence) that it can be reasonably approximated 
with a low-order polynomial in label space.
\item The training set (Section~\ref{sec:the-training-set}) has mean accurate labels
for most, but not all stars. That is to say that we do not assume that \emph{every} 
label in the training is accurate. We can afford to have a small fraction of 
inaccurate labels; few obvious misclassifications in the training set are affordable.
\item We assume that the training data are similar (in spectra) to the test data 
where they overlap in label space, and we assume that the training data spans enough
of the label space to capture the variation in the test-set spectra.
\end{itemize}

@@

\noindent{}where $y_{jn}$ is the pseudo-continuum-normalized flux for star $n$ at wavelength pixel
$j$, $\vecv(\ell_n)$ is the vectorizing function that takes as input the $K$ labels
$\ell_n$ for star $n$ and outputs functions of those labels as a vector of length
$D>K$, $\vectheta_j$ is a vector of length $D$ of parameters influencing the model at
wavelength pixel $j$, and $e_{jn}$ is the residual (noise).  Here we will only consider
vectorizing functions with second-order polynomial expansions (e.g., $\teff^2$, see Sections 
\ref{sec:a-simple-model}-\ref{sec:evolved-star-model}).  The noise values $e_{jn}$ can 
be considered to be drawn from a Gaussian distribution with zero mean and variance 
$\sigma_{jn}^2 + s_j^2$, where $\sigma_{jn}^2$ is the variance in flux $y_{jn}$ and 
$s_j^2$ describes the excess variance at the $j$-th wavelength pixel.

@@

We sought to construct a training set of stars across the main-sequence, the
sub-giant branch, and the red giant branch.  We required stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of O, Mg, Si, Ca, Al, Fe, and Ni.  This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \ges\ survey includes 
giant and main-sequence stars, but only 142 overlap with \rave, which is too small to
be a useful training set for our purposes.  The thirteenth data release from the 
\project{Sloan Digital Sky Survey} \citep{sloan_dr13} includes labels for \apogee\ stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the \apogee\ main-sequence labels suffered from significant systematic effects.  
A flat, then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars incorrectly sit above an isochrone in a classical Hertzsprung-Russell diagram).
If we consider lower-resolution studies as potential training sets, there are 2,369
stars that overlap with \lamost\ --- of which 2,213 have positive S/N ratios in the 
$g$-band (\texttt{snrg}).  However, the labels are expectedly less precise given the
lower resolution, there are no elemental abundances available for the main-sequence 
stars\footnote{Abundance information is available for \lamost\ stars from \citet{Ho_2016},
but that sample contains only giant stars.}, and the \lamost\ lower main-sequence suffers
from the same systematic effects seen in \apogee\ data.

These constraints forced us to construct a heterogeneous training set.  Given previous
successes in transferring high S/N ratio labels from \apogee\ \citep{Ness_2015,
Ness_2016,Ho_2016,Casey_2016b}, we chose to use the 1,355 stars in the \apogee---\rave\ 
overlap sample for giant star labels in the training set.  Of these, about 900 are 
giants according to \apogee.  From this sample we selected stars to have: 
determinations in all abundances of interest ([X/H] $> -5$ for O, Mg, Al, Si, Ca, Fe, 
and Ni); S/N ratios of $>$200~pixel$^{-1}$ in \apogee\ and $>$25~pixel$^{-1}$ in \rave; 
and we further required that \aspcap\ did not report any peculiar flags 
(\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 536 stars along the giant 
branch, with metallicities ranging from $[{\rm Fe/H}] = -1.79$ to 0.26.  Intermediate 
tests with globular cluster members showed that the metallicity range of the training 
set needed to extend at least below $[{\rm Fe/H}] \lesssim -2$ in order for our catalog 
to be practically useful.  Without additional metal-poor stars, the lowest metallicity
labels reported by our model would be around $[{\rm Fe/H}] \approx -2$, even for well
studied stars with $[{\rm Fe/H}] \sim -4$ (e.g., CD~38-245).  For this reason we
supplemented our sample of \apogee\ giant stars with 176 known metal-poor giant stars 
observed by \rave.  The effective temperature $\teff$, surface gravity $\logg$ and
iron abundance [Fe/H] labels were adopted from \citet{Fulbright_2010, Ruchti_2011}.
For this sample of metal-poor stars, we assumed that the elemental abundances of O, 
Mg, Al, Si, Ca, and Ni followed typical trends of Galactic chemical evolution: we
asserted $[{\rm Mg/Fe}] = +0.4$, $[{\rm O/Fe}] = +0.4$, $[{\rm Al/Fe}] = -0.5$, 
$[{\rm Ca/Fe}] = +0.4$, $[{\rm Si/Fe}] = +0.4$, and $[{\rm Ni/Fe}] = -0.25$.  We stress
that this decision is made solely to ensure that our overall metallicity scale reflects
that of the \rave\ survey, down to $[{\rm Fe/H}] \sim -4$.  Indeed, it is likely that 
for most of these elements, these abundances cannot be measured from \rave\ spectra for 
ultra metal-poor stars: the atomic transitions in the \rave\ spectral region are simply
too weak.  For this reason, our adopted abundances for these very metal-poor stars
represent an `anchor point' in order to ensure our overall metallicity scale is
correct.  We do not recommend the use of our individual abundance labels at 
$[{\rm Fe/H}] \sim -4$.  We discuss this issue in more detail in Section~\ref{sec:discussion}.

Assembling a suitable training set for the main-sequence and sub-giant branch was less
trivial.  There are no spectroscopic studies that extend the range of stellar types we 
are interested in (e.g., FGKM-type stars), and which have a large enough sample size 
that overlaps with \rave.  Moreover, most of the spectroscopic studies we considered also 
showed a flat lower main-sequence, a systematic consequence of the analysis method adopted 
\citep[see][for discussion on this issue]{Bensby_2014}.  For these reasons we chose to make 
use of the \epic\ catalog \citep{Huber_2016} for the training set labels on the 
main-sequence and sub-giant branch.  The \epic\ catalog follows from the successful
\project{Kepler} input catalog \citep{Brown_2011}, and provides probabilistic stellar 
classifications for 138,600 stars in the \project{K2} fields based on the 
astrometric, asteroseismic, photometric, and spectroscopic information available for
every star.  There are 4,611 stars that overlap between \epic\ and \rave.

\epic\ differs from the \project{Kepler} input catalog because \epic\ does not 
benefit from having narrow-band $DDO_{51}$ photometry to aid dwarf/giant 
classification.  Despite this limitation, the labels in the \epic\ catalog have 
already been shown to be accurate and trustworthy \citep{Huber_2016}.  However, 
when the posteriors are wide (i.e., the quoted confidence intervals are large) 
due to limited information available, it is possible that a star has been 
misclassified.  This is most prevalent for sub-giants, where \citet{Huber_2016} 
note that $\approx55-70$\% of sub-giants are misclassified as dwarfs.  The 
probability of misclassification is usually quantified in the uncertainties given
for each star; most dwarfs that have a higher possibility of being sub-giants have
large confidence intervals.  Therefore, requiring low uncertainties will decrease 
the total sample size, but in practice it removes most misclassifications.  The 
situation is far more favourable for dwarfs and giants.  Only $1-4$\% of giant 
stars are misclassified as dwarfs, and about 7\% of dwarfs are misclassified as 
giants.  To summarise, the \epic\ labels with narrow confidence intervals are 
usually of high fidelity, and given that we have spectra, we can identify any
spurious misclassifications.

We sought to have a small overlap between our giant and main-sequence star training
sets.  Most of our giant training set is encapsulated within $0 < \logg < 3.5$, 
however there is a sparse sampling of stars reaching to $\logg \approx 4$.  We
required $\logg > 3.5$ for the \epic\ main-sequence/sub-giant star training set,
allowing for $\approx0.5$~dex of overlap between the two training sets.  We further
employed the following quality constraints on the \epic\ catalog: the upper and lower 
confidence intervals in $\teff$ must be below 150~K; the upper and lower confidence 
intervals in $\logg$ must be less than 0.15~dex; the $S/N$ of the \rave\ spectra must
exceed 30~pixel$^{-1}$; and $\teff \leqslant 6750$~K.  Unfortunately these strict 
constraints removed most metal-poor stars, which we later found to cause the test 
labels to have under-predicted abundances for dwarfs of low metallicity.  For this 
reason we relaxed (ignored) those quality constraints for stars with 
$[{\rm Fe/H}] < -1$, and included an additional 12 turn-off stars with 
$-1.6 \gtrsim [{\rm Fe/H}] \gtrsim -2.1$ from \citet{Ruchti_2011}.  After training
a model based on main-sequence and giant stars (Section \ref{sec:the-model}), we found 
we could identify misclassifications by leave-one-out cross-validation.  However, we 
chose not to do this because the number of likely misclassifications in the training
set was negligible ($\approx1$\%), and the improvement in main-sequence test set labels
was minimal.  The distilled sample of the \rave--\epic\ overlap catalog contains 595 
stars (583 of 4,611 from \epic).  The full training set for each model (see next 
sections) is shown in Figure~\ref{fig:training-set-hrd}.

\subsection{A 3-label model ($\teff$, $\logg$, $\feh$) for all stars}
\label{sec:a-simple-model}

We have constructed a justified training set for stars across the main-sequence, sub-giant,
and red giant branch.  However the lack of overlap between \rave\ and other works have
resulted in a somewhat peculiar situation.  Detailed abundances are available from \apogee\
for all giant stars in our sample, but only imprecise metallicities are available from
\epic\ for stars on the main-sequence and the sub-giant branch.  Here we will construct 
a simple model for \emph{all} stars that only makes use of three labels ($\teff$, $\logg$, 
[Fe/H]), before we outline how we derive abundances for giant branch stars.  The complexity
for this model will be quadratic ($\teff^2$ is the highest term), where the vectorizer 
$\vecv(\ell_n)$ expands as,
\begin{eqnarray}\label{eq:vectorizer-three-label}
\vecv(\ell_n) \rightarrow \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\,T_{{\rm eff},n}, \feh_n\,T_{{\rm eff},n}, \logg_n^2, \feh_n\,\logg_n, \feh_n^2\right]
\end{eqnarray}

@@

We chose to adopt separate models for the main-sequence and the red giant branch rather
than switch to a single model with higher complexity.  This choice allowed us to derive
stellar parameters for stars on the main-sequence and sub-giant branch, as well as 
detailed elemental abundances for red giant branch stars.  However, adopting two separate
models introduces the challenge of how to combine the results from two models, or how to
assign one star as `belonging' to a single model.  In Section~\ref{sec:joining-the-models}
we describe how we will use the 3-label model of the main-sequence and giant branch
(introduced in this section) to discriminate between results from a 3-label main-sequence 
model in Section \ref{sec:unevolved-star-model} and a 9-label giant star model in Section 
\ref{sec:evolved-star-model}.

@@

Similar to Sections \ref{sec:a-simple-model} and \ref{sec:unevolved-star-model},
we used a quadratic vectorizer for the giant star model.  Here the terms are 
expanded in the same way as equation \ref{eq:vectorizer-three-label}, only
with nine labels instead of three.  We set the regularization hyperparameter $\Lambda$
in the same way described in Section \ref{sec:unevolved-star-model}, using the same 30 trials of $\Lambda$.
The results are shown in Figure \ref{fig:set-hyperparameters}, where
again the enveloped region represents the minimum and maximum change in RMS label
deviation with respect to the unregularized case.  At the point of maximum improvement 
near $\Lambda = 0.13$ $(\approx10^{-0.9})$, the RMS in all nine labels has decreased by up to $30$\%,
with all labels showing an improvement $>5$\%, and the mean improvement over all labels is about 10\%.
Near $\Lambda \approx 10^{-0.9}$ to $10^{-0.3}$, the regularization also produces a sparser matrix
of $\vectheta$, with $\approx20$\% more terms (mostly cross-terms) having zero-valued entries.
Based on the increased model sparsity and decreasing RMS deviation in the labels, 
we adopt $\Lambda = 0.57$ ($10^{-0.25}$) for the giant star model.  The bias in 
labels from a regularized model with $\Lambda = 0.57$ is negligible: $-0.3$~K in 
$\teff$, and $<$0.007~dex in magnitude for $\logg$ and all seven elements.  The RMS
at this regularization strength is 69~K in $\teff$, 0.18~dex in $\logg$, and varies
between $0.07-0.09$~dex depending on the elemental abundance.

@@

\subsection{Deriving joint estimates from all models}
\label{sec:joining-the-models}

@@

In Figure \ref{fig:joint-model-differences} we show the differences in effective 
temperature $\teff$ and surface gravity $\logg$ between: the main-sequence model 
(Section~\ref{sec:unevolved-star-model}) and the simplest model (Section~
\ref{sec:a-simple-model}); and the differences between the red giant branch model 
(Section~\ref{sec:evolved-star-model}) and the simplest model (Section 
\ref{sec:a-simple-model}).  We have scaled the differences in $\teff$ and $\logg$ 
to make the central peak near $(0, 0)$ to be approximately isotropic by setting
	$\delta_{T_{\rm eff}} = 90$~K for the main-sequence model, 
	$\delta_{T_{\rm eff}} = 50$~K for the giant model, and 
	$\delta_{\log{g}} = 0.15$~dex for both models.
It is important to note that these $\delta$ values do not represent any kind of 
intrinsic uncertainty or precision: they are merely normalization factors.  Adopting
substantially different scaling factors would produce in clear inconsistencies in our
results (e.g., sub-giants being misclassified as giants).  Therefore we chose these 
factors empirically to make the distributions in Figure \ref{fig:joint-model-differences}
approximately isotropic, and to some extent, comparable.  In Figure 
\ref{fig:joint-model-differences} the stars within the peak at $(0, 0)$ represent 
objects where the simple model and the comparison model both report similar labels.  
The artefacts seen in the Hertzsprung-Russell diagrams in Figure \ref{fig:test-set-density} 
are also present in Figure \ref{fig:joint-model-differences} as over-densities far away
from the central peak.  Therefore, we can adopt the scaled distance in labels 
$\teff$ and $\logg$ from the simplest model to the main-sequence model $d_{ms}$,
\begin{eqnarray}
	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{joint}}{\delta_{\log{g},ms}}\right)^2 \quad ,
\end{eqnarray}

\noindent{}and the scaled distance from the simplest model to the giant model $d_{giant}$,
\begin{eqnarray}
	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},joint}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{joint}}{\delta_{\log{g},giant}}\right)^2  \quad ,
\label{eq:d_giant}
\end{eqnarray}

>>>>

\begin{abstract}
The Milky Way is a powerful laboratory for understanding galaxy formation,
as it can provide orbits, ages, atmospheric parameters, as well as chemical 
abundances for vast sets of individual stars.  These inferences require 
both astrometric and spectroscopic data.  Indeed, in order to be maximally
useful for chemo-dynamic studies, the Tycho-Gaia Astrometric Solution 
(\tgas) sample in the first \gaia\ data release requires a spectroscopic 
complement that includes radial velocities, stellar atmospheric parameters, and 
elemental abundances.  Among existing spectroscopic samples, the RAdial 
Velocity Experiment (\rave) survey has the largest overlap with \tgas: 
up to about 264,000 stars.  We present a data-driven re-analysis of \rave\ 
spectra using \thecannon, and we provide more precise and accurate stellar
parameters and abundances than available in previous \rave\ data releases. 
We constructed our model using high-fidelity \apogee\ stellar parameters 
and abundances for red giant branch stars that overlap with \rave, and 
stellar parameters from the \project{K2}/\project{Ecliptic Pole Input 
Catalog} for main-sequence and sub-giant branch stars.  We derive, and 
validate, improved effective temperature $\teff$, surface gravity $\logg$, 
and chemical abundances of up to seven elements (O, Mg, Al, Si, Ca, Fe, Ni) 
for \Nstarsqc\ stars.  In total we report \Nabundances\ elemental 
abundances, with a typical precision of 0.07~dex.  The \unrave\ catalog 
presented here, in conjunction with \tgas, represents the most powerful 
data set for chemo-dynamic analysis at the dawn of broadly available 
\gaia\ data.
\end{abstract}

@@

\begin{itemize}
\item We assume that any fibre- and time-dependent variations in spectral
resolution in the \rave\ spectra are negligible.
\item The \rave\ noise variances are approximately correct, independent between
pixels, and normally distributed.
\item We assume that the normalization procedure employed by the \rave\ pipeline
is invariant with respect to the labels we seek to measure (e.g., $\teff$, $\logg$,
or [Fe/H]), and invariant with respect to the S/N ratio.  In other words, we assume
that the normalization procedure does not produce different results for high S/N
spectra compared to low S/N spectra, nor does the normalization procedure vary 
non-linearly with respect to stellar parameters (e.g., [Fe/H]).
\item We assume that stars with similar labels ($\teff$, $\logg$, and abundances)
have similar spectra.
\item A stellar spectrum is a smooth function of the label values for that star,
and we assume that the function is smooth enough within a sub-space of the labels
(e.g., the giant branch or the main-sequence) that it can be reasonably approximated 
with a low-order polynomial in label space.
\item The training set (Section~\ref{sec:the-training-set}) has mean accurate labels
for most, but not all stars. That is to say that we do not assume that \emph{every} 
label in the training is accurate. We can afford to have a small fraction of 
inaccurate labels; a few obvious misclassifications in the training set are affordable.
\item We assume that the training data are similar (in spectra) to the test data 
where they overlap in label space, and we assume that the training data spans enough
of the label space to capture the variation in the test-set spectra.
\end{itemize}

@@

\noindent{}where $y_{jn}$ is the pseudo-continuum-normalized flux for star $n$ at wavelength pixel
$j$, $\vecv(\ell_n)$ is the vectorizing function that takes as input the $K$ labels
$\ell_n$ for star $n$ and outputs functions of those labels as a vector of length
$D>K$, $\vectheta_j$ is a vector of length $D$ of parameters influencing the model at
wavelength pixel $j$, and $e_{jn}$ is the residual (noise).  Here we will only consider
vectorizing functions with second-order polynomial expansions (e.g., $\teff^2$, see Sections 
\ref{sec:a-simple-model}--\ref{sec:evolved-star-model}).  The noise values $e_{jn}$ can 
be considered to be drawn from a Gaussian distribution with zero mean and variance 
$\sigma_{jn}^2 + s_j^2$, where $\sigma_{jn}^2$ is the variance in flux $y_{jn}$ and 
$s_j^2$ describes the excess variance at the $j$-th wavelength pixel.

@@

We sought to construct a training set of stars across the main-sequence, the
sub-giant branch, and the red giant branch.  We required stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of O, Mg, Si, Ca, Al, Fe, and Ni.  This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \ges\ survey includes 
giant and main-sequence stars, but only 142 overlap with \rave, which is too small to
be a useful training set for our purposes.  The thirteenth data release from the 
\project{Sloan Digital Sky Survey} \citep{sloan_dr13} includes labels for \apogee\ stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the \apogee\ main-sequence labels suffered from significant systematic effects.  
A flat, then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars incorrectly sit above an isochrone in a classical Hertzsprung-Russell diagram).
If we consider lower-resolution studies as potential training sets, there are 2,369
stars that overlap with \lamost\ --- of which 2,213 have positive S/N ratios in the 
$g$-band (\texttt{snrg}).  However, the labels are expectedly less precise given the
lower resolution, there are no elemental abundances available for the main-sequence 
stars\footnote{Abundance information is available for \lamost\ stars from \citet{Ho_2016},
but that sample contains only giant stars.}, and the \lamost\ lower main-sequence suffers
from the same systematic effects seen in the \apogee\ data.

These constraints forced us to construct a heterogeneous training set.  Given previous
successes in transferring high S/N ratio labels from \apogee\ \citep{Ness_2015,
Ness_2016,Ho_2016,Casey_2016b}, we chose to use the 1,355 stars in the \apogee---\rave\ 
overlap sample for giant star labels in the training set.  Of these, about 900 are 
giants according to \apogee.  From this sample we selected stars to have: 
determinations in all abundances of interest ([X/H] $> -5$ for O, Mg, Al, Si, Ca, Fe, 
and Ni); S/N ratios of $>$200~pixel$^{-1}$ in \apogee\ and $>$25~pixel$^{-1}$ in \rave; 
and we further required that the \aspcap\ did not report any peculiar flags 
(\texttt{ASPCAPFLAG = 0}).  These restrictions left us with 536 stars along the giant 
branch, with metallicities ranging from $[{\rm Fe/H}] = -1.79$ to 0.26.  Intermediate 
tests with globular cluster members showed that the metallicity range of the training 
set needed to extend at least below $[{\rm Fe/H}] \lesssim -2$ in order for our catalog 
to be practically useful.  Without additional metal-poor stars, the lowest metallicity
labels reported by our model would be around $[{\rm Fe/H}] \approx -2$, even for well
studied stars with $[{\rm Fe/H}] \sim -4$ (e.g., CD~38-245).  For this reason we
supplemented our sample of \apogee\ giant stars with 176 known metal-poor giant stars 
observed by \rave.  The effective temperature $\teff$, surface gravity $\logg$ and
iron abundance [Fe/H] labels were adopted from \citet{Fulbright_2010, Ruchti_2011}.
For this sample of metal-poor stars, we assumed that the elemental abundances of O, 
Mg, Al, Si, Ca, and Ni followed typical trends of Galactic chemical evolution: we
asserted $[{\rm Mg/Fe}] = +0.4$, $[{\rm O/Fe}] = +0.4$, $[{\rm Al/Fe}] = -0.5$, 
$[{\rm Ca/Fe}] = +0.4$, $[{\rm Si/Fe}] = +0.4$, and $[{\rm Ni/Fe}] = -0.25$.  We stress
that this decision is made solely to ensure that our overall metallicity scale reflects
that of the \rave\ survey, down to $[{\rm Fe/H}] \sim -4$.  Indeed, it is likely that 
for most of these elements, these abundances cannot be measured from \rave\ spectra for 
ultra metal-poor stars: the atomic transitions in the \rave\ spectral region are simply
too weak to influence the spectrum.  For this reason, our adopted abundances for these
very metal-poor stars represent an `anchor point' in order to ensure our overall 
metallicity scale is correct.  We do not recommend the use of our individual abundance
labels at $[{\rm Fe/H}] \sim -4$.  We discuss this issue in more detail in Section~\ref{sec:discussion}.

Assembling a suitable training set for the main-sequence and sub-giant branch was less
trivial.  There are no spectroscopic studies that extend the range of stellar types we 
are interested in (e.g., FGKM-type stars), and which also have a large enough sample size 
that overlaps with \rave.  Moreover, most of the spectroscopic studies we considered also 
showed a flat lower main-sequence, a systematic consequence of the analysis method adopted 
\citep[see][for discussion on this issue]{Bensby_2014}.  For these reasons we chose to make 
use of the \epic\ catalog \citep{Huber_2016} for the training set labels on the 
main-sequence and sub-giant branch.  The \epic\ catalog follows from the successful
\project{Kepler} input catalog \citep{Brown_2011}, and provides probabilistic stellar 
classifications for 138,600 stars in the \project{K2} fields based on the 
astrometric, asteroseismic, photometric, and spectroscopic information available for
every star.  There are 4,611 stars that overlap between \epic\ and \rave.

\epic\ differs from the \project{Kepler} input catalog because \epic\ does not 
benefit from having narrow-band $DDO_{51}$ photometry to aid dwarf/giant 
classification.  Despite this limitation, the labels in the \epic\ catalog have 
already been shown to be accurate and trustworthy \citep{Huber_2016}.  However, 
when the posteriors are wide (i.e., the quoted confidence intervals are large) 
due to limited information available, it is possible that a star has been 
misclassified.  This is most prevalent for sub-giants, where \citet{Huber_2016} 
note that $\approx55-70$\% of sub-giants are misclassified as dwarfs.  The 
probability of misclassification is usually quantified in the uncertainties given
for each star; most dwarfs that have a higher possibility of being sub-giants have
large confidence intervals.  Therefore, requiring low uncertainties will decrease 
the total sample size, but in practice it removes most misclassifications.  The 
situation is far more favourable for dwarfs and giants.  Only 1--4\% of giant 
stars are misclassified as dwarfs, and about 7\% of dwarfs are misclassified as 
giants.  To summarise, the \epic\ labels with narrow confidence intervals are 
usually of high fidelity, and given that we have spectra, we can identify any
obvious misclassifications.

We sought to have a small overlap between our giant and main-sequence star training
sets.  Most of our giant training set is encapsulated within $0 < \logg < 3.5$, 
however there is a sparse sampling of stars reaching to $\logg \approx 4$.  We
required $\logg > 3.5$ for the \epic\ main-sequence/sub-giant star training set,
allowing for $\approx0.5$~dex of overlap between the two training sets.  We further
employed the following quality constraints on the \epic\ catalog: the upper and lower 
confidence intervals in $\teff$ must be below 150~K; the upper and lower confidence 
intervals in $\logg$ must be less than 0.15~dex; the S/N of the \rave\ spectra must
exceed 30~pixel$^{-1}$; and $\teff \leqslant 6750$~K.  Unfortunately these strict 
constraints removed most metal-poor stars, which we later found to cause the test 
labels to have under-predicted abundances for dwarfs of low metallicity.  For this 
reason we relaxed (ignored) those quality constraints for stars with 
$[{\rm Fe/H}] < -1$, and included an additional 12 turn-off stars with 
$-1.6 \gtrsim [{\rm Fe/H}] \gtrsim -2.1$ from \citet{Ruchti_2011}.  After training
a model based on main-sequence and giant stars (Section \ref{sec:the-model}), we found 
we could identify misclassifications by leave-one-out cross-validation.  However, we 
chose not to do this because the number of likely misclassifications in the training
set was negligible ($\approx$1\%), and the improvement in main-sequence test set labels
was minimal.  The distilled sample of the \rave--\epic\ overlap catalog contains 595 
stars (583 of 4,611 from \epic).  The full training set for each model (see next 
sections) is shown in Figure~\ref{fig:training-set-hrd}.

\subsection{The simple model: a 3-label model ($\teff$, $\logg$, $\feh$) for all stars}
\label{sec:a-simple-model}

We have constructed a justified training set for stars across the main-sequence, sub-giant,
and red giant branch.  However the lack of overlap between \rave\ and other works have
resulted in a somewhat peculiar situation.  Detailed abundances are available from \apogee\
for all giant stars in our sample, however only imprecise (but accurate on expectation)  
metallicities are available from \epic\ for stars on the main-sequence and the sub-giant 
branch.  Here we will construct a simple model for \emph{all} stars that only makes use 
of three labels ($\teff$, $\logg$, [Fe/H]), before we outline how we derive abundances 
for giant branch stars.  The complexity for this model will be quadratic ($\teff^2$
is the highest term), where the vectorizer $\vecv(\ell_n)$ expands as,
\begin{eqnarray}\label{eq:vectorizer-three-label}
\vecv(\ell_n) \rightarrow \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\,T_{{\rm eff},n}, \feh_n\,T_{{\rm eff},n}, \logg_n^2, \feh_n\,\logg_n, \feh_n^2\right]
\end{eqnarray}

@@

We chose to adopt separate models for the main-sequence and the red giant branch rather
than switch to a single model with higher complexity.  This choice allowed us to derive
stellar parameters for stars on the main-sequence and sub-giant branch, as well as 
detailed elemental abundances for red giant branch stars.  However, adopting two separate
models introduces the challenge of how to combine the results from two models, or how to
assign one star as `belonging' to a single model.  In Section~\ref{sec:joining-the-models}
we describe how we will use the simple model introduced in this section to discriminate
between results from a 3-label main-sequence model in Section \ref{sec:unevolved-star-model} 
and a 9-label giant star model in Section \ref{sec:evolved-star-model}.

@@

Similar to Sections \ref{sec:a-simple-model} and \ref{sec:unevolved-star-model},
we used a quadratic vectorizer for the giant star model.  Here the terms are 
expanded in the same way as equation \ref{eq:vectorizer-three-label}, only
with nine labels instead of three.  We set the regularization hyperparameter $\Lambda$
in the same way described in Section \ref{sec:unevolved-star-model}, using the same 30 trials of $\Lambda$.
The results are shown in Figure \ref{fig:set-hyperparameters}, where
again the enveloped region represents the minimum and maximum change in RMS label
deviation with respect to the unregularized case.  At the point of maximum improvement 
near $\Lambda = 0.13$ $(\approx10^{-0.9})$, the RMS in all nine labels has decreased by up to $30$\%,
with all labels showing an improvement $>5$\%, and the mean improvement over all labels is about 10\%.
Near $\Lambda \approx 10^{-0.9}$ to $10^{-0.3}$, the regularization also produces a sparser matrix
of $\vectheta$, with $\approx20$\% more terms (mostly cross-terms) having zero-valued entries.
Based on the increased model sparsity and decreasing RMS deviation in the labels, 
we adopt $\Lambda = 0.57$ ($10^{-0.25}$) for the giant star model.  The bias in 
labels from a regularized model with $\Lambda = 0.57$ is negligible: $-0.3$~K in 
$\teff$, and $<$0.007~dex in magnitude for $\logg$ and all seven elemental
abundances.  The RMS at this regularization strength is 69~K in $\teff$, 0.18~dex 
in $\logg$, and varies between $0.07-0.09$~dex depending on the elemental abundance.

@@

\subsection{Deriving joint estimates from multiple models}
\label{sec:joining-the-models}

@@

In Figure \ref{fig:joint-model-differences} we show the differences in effective 
temperature $\teff$ and surface gravity $\logg$ between: the main-sequence model 
(Section~\ref{sec:unevolved-star-model}) and the simple model (Section 
\ref{sec:a-simple-model}); and the differences between the red giant branch model 
(Section~\ref{sec:evolved-star-model}) and the simple model (Section 
\ref{sec:a-simple-model}).  We have scaled the differences in $\teff$ and $\logg$ 
to make the central peak near $(0, 0)$ to be approximately isotropic by setting
	$\delta_{T_{\rm eff}} = 90$~K for the main-sequence model, 
	$\delta_{T_{\rm eff}} = 50$~K for the giant model, and 
	$\delta_{\log{g}} = 0.15$~dex for both models.
It is important to note that these $\delta$ values do not represent any kind of 
intrinsic uncertainty or precision: they are merely normalization factors.  Adopting
substantially different scaling factors would produce in clear inconsistencies in our
results (e.g., sub-giants being misclassified as giants).  Therefore we chose these 
factors empirically to make the distributions in Figure \ref{fig:joint-model-differences}
approximately isotropic, and to some extent, comparable.  In Figure 
\ref{fig:joint-model-differences} the stars within the peak at $(0, 0)$ represent 
objects where the simple model and the comparison model both report similar labels.  
The artefacts seen in the Hertzsprung-Russell diagrams in Figure \ref{fig:test-set-density} 
are also present in Figure \ref{fig:joint-model-differences} as over-densities far away
from the central peak.  Therefore, we can adopt the scaled distance in labels 
$\teff$ and $\logg$ from the simple model to the main-sequence model $d_{ms}$,
\begin{eqnarray}
	d_{ms} & = & \left(\frac{T_{{\rm eff},ms} - T_{{\rm eff},simple}}{\delta_{T_{{\rm eff},ms}}}\right)^2 + \left(\frac{\log{g}_{ms} - \log{g}_{simple}}{\delta_{\log{g},ms}}\right)^2 \quad ,
\end{eqnarray}

\noindent{}and the scaled distance from the simple model to the giant model $d_{giant}$,
\begin{eqnarray}
	d_{giant} & = & \left(\frac{T_{{\rm eff},giant} - T_{{\rm eff},simple}}{\delta_{T_{{\rm eff},giant}}}\right)^2 + \left(\frac{\log{g}_{giant} - \log{g}_{simple}}{\delta_{\log{g},giant}}\right)^2  \quad ,
\label{eq:d_giant}
\end{eqnarray}
